---
output:
  word_document: default
  pdf_document:
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = F)
library(here)
library(tidyverse)
library(spdep)
library(knitr)
library(kableExtra)
library(texreg)
```

## Appendix

In this appendix we present the complete estimation results for the models 
relating different definitions of access to physical activity (in Table
\ref{tab:pa-fullmodels}) and to obesity (in Table \ref{tab:ob-fullmodels}). In
each case we also present a base model with no accessibility statistics for
comparison.

```{r}
pa_models <- read_rds(here("data/pa_models.rds"))
obesity_models <- read_rds(here("data/obesity_models.rds"))
```

```{r}
base_termnames <- c(
  "(Intercept)", "log(Density)", "log(Income)", "Fulltime", "College-educated",  "Single Adults",
  "Youth (0-17)", "Young adults (18-34)", "Seniors (65+)", "Black population share",
  "Asian population share", "Hispanic population share", "Other Minorities"
)

cmnames <- c(
  base_termnames, str_c("$\\gamma$: ", base_termnames[-1]), 
  rep("Accessibility", 3), "$\\lambda$: spatial correlation"
)

#coefficient map
cm <- as.list(cmnames)
names(cm) <- c(names(pa_models$Logsum$coefficients), "euc_multi", "walk10TRUE", "$\\lambda$")

# obesity coefficient map
cmonames <- c(
  base_termnames, str_c("$\\gamma$: ", base_termnames[-1]), 
  "Physical Activity", "$\\gamma$: Physical Activity",
  rep("Accessibility", 3),
  "$\\lambda$: spatial correlation"
)
cmo <- as.list(cmonames)
names(cmo) <- c(names(obesity_models$Base$coefficients), 
                "euc_access", "euc_multi", "walk10TRUE",  "$\\lambda$")
```

\begin{table*}
  \caption{\label{tab:pa-fullmodels} Estimated Effect of Accessibility on Physical Activity Rates}
  \begin{center}
  \scalebox{0.8}{
```{r pa-fullmodels, results = "asis"}
texreg(
  pa_models, digits = 4, ci.force = TRUE,
  booktabs = TRUE, use.packages = FALSE, 
  custom.coef.map = cm,
  custom.model.names = c("No Access", "Size and Distance", "Amenities", "10-minute Walk"), 
  table = F,
  include.aic = F, include.lr = F,
  custom.note = "%stars 95\\% confidence interval in brackets.")
```
}\end{center}
\end{table*}


\begin{table*}
  \caption{\label{tab:ob-fullmodels} Estimated Effect of Accessibility on Obesity Rates}
  \begin{center}
  \scalebox{0.8}{
```{r ob-fullmodels, results="asis"}
texreg(
  obesity_models, digits = 4, ci.force = TRUE,
  booktabs = TRUE, use.packages = FALSE, 
  custom.coef.map = cmo,
  custom.model.names = c("No Access", "Size and Distance", "Amenities", "10-minute Walk"), 
  table = F,
  include.aic = F, include.lr = F,
  custom.note = "%stars 95\\% confidence interval in brackets.")
```
}\end{center}
\end{table*}
