---
title: >
  Modeling the impacts of park access on health outcomes: a choice-based 
  accessibility approach
author:
  - name: Gregory Macfarlane
    email: gregmacfarlane@byu.edu
    affiliation: Brigham Young University
    corresponding: gregmacfarlane@byu.edu
  - name: Nico Boyd
    email:
    affiliation: Fehr and Peers 
  - name: John E. Taylor
    email: jetaylor@gatech.edu
    affiliation: Georgia Institute of Technology
  - name: Kari Watkins
    email: kariwatkins@gatech.edu
    affiliation: Georgia Institute of Technology
address:
  - code: Brigham Young University
    address: Civil and Environmental Engineering Department, Brigham Young University 430 Engineering Building, Provo, UT, 84602
  - code: Fehr and Peers
    address: Los Angeles, California
  - code: Georgia Institute of Technology
    address: School of Civil and Environmental Engineering, Georgia Institute of Technology, 755 Atlantic Drive, Atlanta, GA, 30332
abstract: >
  Recent research has underscored the potential for public green spaces to
  influence individual and societal health outcomes, but empirical measurements
  of such influences have yielded mixed results to date, with particular
  disagreement surrounding how access to parks ought to be defined while
  controlling alternate explanations. In this paper we present a comprehensive
  measure of park accessibility drawn from random utility choice theory, which
  avoids arbitrary assertions of proximity while incorporating potentially
  numerous amenities and attributes of both the parks and the population. We
  apply this choice-based accessibility measure to correlate Census tract-level
  obesity and physical activity rate estimates from the Centers for Disease
  Control and Prevention's 500 Cities project with tract-level American
  Community Survey socioeconomic data in New York City, paired with geographic
  parks data from New York City. Controlling for the
  socioeconomic variables and spatially correlated error terms, we show a
  positive and significant relationship between park access and physical
  activity rates, and a clear and significant negative relationship between park
  access and obesity rates. In doing so, this research contributes a more
  comprehensive modeling approach for measuring the impact of park access on
  health, and may improve our understanding of the role  parks and access to
  them can serve in furthering public health objectives.
author_summary: >
  Greg Macfarlane is an assistant professor of Civil Engineering at Brigham
  Young University in Provo, Utah.

output: 
  rticles::elsevier_article:
    includes: 
      in_header: preamble.tex
csl: elsevier-harvard.csl
bibliography: library.bib
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(geojsonio)
library(rgdal)
library(rgeos)
library(maptools)
library(spdep)
library(sf)
library(VGAM)
library(qwraps2)
library(knitr)
library(kableExtra)
library(texreg)
library(here)

options(tigris_use_cache = TRUE)
opts_chunk$set(cache=TRUE, message = FALSE, echo = FALSE, warning = FALSE)
library(leaflet)

knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark=",", drop0trailing = TRUE)
})

```

# Introduction

The United States and other developed nations face an epidemic of obesity and
chronic diseases, including cardiovascular diseases, chronic respiratory
diseases, diabetes, stroke, joint and bone diseases, and cancer. These diseseases
can severely limit the lifespan of affected individuals [@WHO2014], with 
substantial financial costs for treatment borne both privately and socially
[@Finkelstein2009]. Though a moderate amount of regular physical activity
has been established as an effective strategy for reducing and managing obesity
and many of the aforementioned chronic diseases [@CDC2009; @Durstine2013],
a large portion of U.S. adults do not participate in sufficient physical
activity [@Wolf2008].

Historically, most large-scale health promotion efforts focused on
individual-level interventions intended to educate people about healthy
lifestyles and behaviors, touching on topics including diet and exercise.
Over the last few years a new social model of health has evolved, treating health
``as an outcome of the effects of socioeconomic status, culture,
environmental conditions, housing, employment and community influences.''[@Duhl1999, p.7]
In this new paradigm, resources provided via civil infrastructure --- in
particular parks and other public green spaces --- play a crucial role in
promoting and sustaining public health [@Bedimo-Rung2005; @Coutts2008].
After all, it does little to lecture individuals on the importance of exercise
if their community is built in such a way that such exercise is expensive,
unenjoyable, or unsafe.

In spite of the potential for green spaces to influence public health outcomes 
and numerous studies attempting to empirically estimate the strength of these
outcomes, evidence to date has been mixed. A major challenge researchers have
faced is a proliferation of techniques for measuring access to green spaces, of
which many are plausible but lack robust theoretical basis. In this
paper, we present a holistic and flexible measurement for park accessibility
based in activity location choice theory. This choice-based accessibility measure
compares the continuous distance to all parks in the region, weighted against
the sizes of the each park and other assorted amenities.

We apply this measurement to study the link between
park accessibility and attractiveness and Census tract-level aggregate physical
activity participation and obesity rates in New York City, controlling for
spatially correlated unobserved effect and tract-level socioeconomic
characteristics. We find a positive relationship where the least park-accessible
tracts have an expected physical activity participation rate roughly a full
percentage point lower than the most accessible tracts; a traditional
quarter-mile buffer analysis estimates a similar albeit marginally less
significant effect. However, we also find a strong and significant correlation
between the choice-based park accessibility metric and decreased obesity rates
--- again controlling for spatial correlation, socioeconomic attributes, and
physical activity rates. In this case a traditional buffer analysis does not
show significance.

We also demonstrate how the choice-based accessibility metric can potentially be
expanded to account for other measures of a park's attractiveness including its
observed social media activity and the presence of various park amenities.
The paper concludes with a discussion of opportunities for future research.

# Literature Review 
\label{sec:litreview}

Numerous previous studies have found relationships between characteristics of
the built environment and aspects of physical, mental, social, and economic
health. For the purposes of this study, it is clear that parks and other
vegetated areas support outdoor physical activity in ways distinct from other
urban environments [@Giles-Corti2005; @Wells2007].  In spite of these generally
positive findings, the evidence of a specific link between green space and
physical activity is somewhat mixed.

<!-- Rewrite this section: 
  - Cross-metropolitan studies are somewhat clear?
  - Within-metropolitan studies are limited
-->

@Wolf2008 reports that people self-identify as more active and attain a
higher quality of life when a greater portion of their environment is greenspace.
@Larson2016 similarly used self-reported scores on the Gallup-Healthways
Wellbeing Index to evaluate the relationship between different areas of
wellbeing, including physical health, and park quantity, quality, and
accessibility in 44 U.S. cities. While the authors found positive associations
between wellbeing and park quality and accessibility, these relationships were
not statistically significant. A study by @West2012 used park data from
the Trust for Public Land’s 2010 City Park Facts and public health data from the
Behavioral Risk Factor Surveillance System (BRFSS) to examine relationships
between the density of parkland, parkland per capita, and levels of physical
activity and obesity for 67 metropolitan statistical areas in the U.S. The study
found a significant, positive association between park density and physical
activity and a significant, negative association between park density and
obesity. Conversely, @Richardson2012 examined the relationship between
urban greenspace and selected mortality rates, and though the authors did not
find a statistically significant relationship between the quantity of urban
greenspace and mortality caused by lung cancer, diabetes, heart disease or car
accidents, they did find that all-cause mortality was oddly, *higher* in
greener cities. In a meta-analysis of 20 peer reviewed journal articles
exploring the relationship between parks and objectively measured physical
activity, @Bancroft2015 found that five studies reported a significant
positive association between the two, six studies produced mixed results, and
nine studies found no association at all.

While several of the aforementioned studies reported results consistent with the
hypothesis that park access and use confer health benefits stemming from
increased physical activity and attendant decreases in obesity, the large
majority of extant research has been conducted at the level of the city or
metropolitan statistical area. Metropolitan-level analyses do not capture the
within-city variation in accessibility to parks that exist in many cities. Some
cities with large amounts of total greenspace may nevertheless unequally
distribute this space throughout the city, leading to areas with poor access.
Conversely, a city with a smaller overall proportion of greenspace may give all
of its citizens better access to sufficient greenspace to meet their needs or
wants for physical activity. Considering access at the neighborhood level within
a single city may also eliminate some regional or cultural fixed effects
affecting metropolitan-level analyses.

In a study of neighborhoods in New York City, @Stark2014 used
hierarchical linear models to explore relationships between the proportion of
zip codes that was park space, the cleanliness of parks, and the body mass index
of adult residents. The authors found that a higher proportion of park space in
a neighborhood was associated with a lower BMI for its residents, and lower park
cleanliness scores were associated with a higher BMI, suggesting that increasing
the supply of clean, well-maintained parks could positively impact adult BMI.


## Choice-based Accessibility

Consider that an individual is choosing a park for a recreation activity.
If we apply random utility choice theory, specifically the multinomial
logit model [@McFadden1974], the expected probability of individual
$i$ choosing park $j$ from the set of all regional parks $J$ is
\begin{equation}\label{eq:mnl}
P(j | V_{ij}) = \frac{\exp(V_{ij})}{\sum_{p \in J}\exp(V_{ip})}
\end{equation}
where parks are differentiated from each other by their relative measurable
utilities $V_{ij}$. In principle, $V$ may include any measurable attributes of either
the choice maker or the park. In this study we use a linear formulation of
\begin{equation}\label{eq:v_sd}
V_{ij} = size_j\lambda_s + distance_{ij}\lambda_d
\end{equation}
incorporating the size of the park in acres and the distance of the park from
the census tract $i$ in miles. The coefficients $\vect{\lambda}$ can be
estimated from household surveys, though in the absence of a survey we
may assert reasonable values.

A key theoretical outcome of random utility choice models is that the
consumer surplus to individual $i$, or the total weighted value of all
alternatives in individual $i$'s choice set, can
be obtained as the logarithm of the sum in the denominator of the choice
probability equation [@Small1981]:
\begin{equation}\label{eq:cs}
C_i = \ln\left({\sum_{p \in J}\exp(V_{ip})}\right)
\end{equation}
There are several advantages to such a log-sum defined metric relative to
buffer-based accessibility metrics more commonly found in the literature. First,
all individuals are defined as having some access to all parks, rather than an
arbitrary cutoff asserted by the researcher.
This allows for the fact that some people are more or less sensitive to
distances, and that distance is a continuous, non-binary phenomenon.
It defies reason to assume a person living 1.1 miles
from a park has functionally different accessibility than someone living at 0.9 miles.
Second, the random utility formulation
allows the researcher to include any attribute of the park as part of its utility
specification; in this case, we consider the size of the park weighted against its
distance as an element of relative accessibility. This suggests that
not all parks are equal, and that a large park such as Central Park in Manhattan
may provide health and activity benefits over a much larger area than a smaller
community square.

Note that this formulation is an extension of gravity-based accessibility
statistics, in the same way that the gravity-based trip distribution model
is a specific case of
a destination-choice model [@Daly1982]. But the extension is meaningful and
allows us to consider more attributes of the park or the individual beyond size
and impedance. For example, we can aggregate the number of tweets emanating from
a park as a measure of the park's attractiveness beyond its size. The
utility for the choice maker from Equation \ref{eq:v_sd} then becomes
\begin{equation}\label{eq:v_sdt}
V_{ij} = size_j\lambda_s + distance_{ij}\lambda_d + tweets_j\lambda_t
\end{equation}

In spite of its flexibility and basis in choice theory, accessibility measures
based in log-sums have not received much application in the accessibility
literature compared with distance-based or even gravity-based measures.
@Zhang2011 present a metric with a similar functional form, though the
authors arrive at it from a motivation other than destination choice theory.
Log-sums are commonly used, however, in alternatives analyses of transit
infrastructure improvements [@DeJong2007]. A reason for this is likely that
a regional travel demand model is available to the analysts, thus making
calibrated and multimodal logsums readily available [@Geurs2010].

# Empirical Application
In this section, we develop a model with data for New York City, where
we compute a holistic logsum-based accessibility to parks measure and model the
relationship between this measure and physical activity rates, controlling for
spatial effects and socioeconomic factors. We subsequently model the effect
of the accessibility metric on obesity rates, accounting for physical activity and
the other controls. Finally, we model the relationship between a modified logsum
measure considering geolocated Twitter data and physical activity and obesity rates.

## Data

This study uses data available to the public from a variety of federal and state
data agencies. The datasets, as well as the analysis code, are available on GitHub
at \url{https://github.com/gregmacfarlane/parks_access}.

```{r tracts}
# built in 01_health_data.Rmd and 03_logsums.Rmd
tracts <- read_rds(here("data/tracts_access.rds"))
```

The Centers for Disease Control and Prevention makes small-area estimates on
key health indicators available through its 500 Cities data program [@CDC5002016].
The indicators are multilevel aggregations and imputations of BRFSS responses
[@Wang2018; @Wang2017], and have been recently used to study the tract-level
link between gentrification and urban health [@Gibbons2018]. We use two
indicators as our dependent variables in this study: the share of adults
in a Census tract who are obese, and the share of adults who participate in no
leisure-time physical activity. To improve clarity in our interpretation, we use
the complement of the second variable --- the share of tract adults who
participate in *some* physical activity --- even if the amount is not sufficient
to affect overall health. Both indicators are estimated for the year 2016.

To the health data, we join socioeconomic data collected through the Census
Bureau API via the \texttt{tidycensus} package for R [@Walker2019]. The
primary dataset is a geographic polygons layer of Census tracts in the five
boroughs of New York City. We append to each Census tract relevant
sociodemographic data for each tract from the American Community Survey
2013-2017 5-year estimates. For a small handful of tracts in our sample, Census
supressed the median income estimate; these appear to be primarily wealthy
tracts and in almost all cases the CDC estimates of obesity and physical
activity are missing as well. After removing these tracts from the estimation
dataset, we have `r nrow(tracts)` complete cases. Table \ref{tab:tractsdata}
presents key descriptive statistics for these data.

```{r parks}
# This data is created in 02_parks_data.Rmd
parks <- read_rds(here("data/open_spaces.rds"))
```

In a destination choice framework, the tracts represent the ``origins'' and
``destinations'' are parks and green spaces in New York City.
We retrieved a polygons layer of public parks and greenspaces within New York City’s
municipal boundaries and checked it for accuracy and relevance [@nycparks].
Upon inspection, we removed several facilities that do not qualify as
publicly accessible green space, such as Yankee Stadium, Citi Field, and their
surrounding parking lots. We also removed parks of less than half an acre in
size, as these appear to be predominantly planted medians rather than legitimate
public green space. We also consolidated individual geographic polygons
comprising a single facility --- as is the case in Flushing Meadows --- and
eliminated sub-facilities such as tennis courts or baseball fields included within
larger park facilities. Instead of these sub-facilities, we created 
variables for each park indicating the presence of sports courts, playgrounds,
and trails.

To this dataset of parks and regularized green spaces, we add a polygons layer of
cemeteries in New York that are open to the public. Cemeteries are important
green spaces that can be used for many types of physical activity. These
operations leave us with `r nrow(parks)` distinct parks. For each park and cemetery
we generate calculate the size of the park in acres. Using a Twitter application
programming interface (API) implemented with the Python package Tweepy,  we also
collected and stored tweets containing geotags of precise geographic coordinates
located within the boundaries of the parks; from this information,
we were able to segregate tweets that were generated within a park in September
2014. This geolocated twitter activity could provide an additional point of information
on the degree to which individual parks are actually used [@Wang2016].

```{r tractsdata, results="asis", message=FALSE, warning=FALSE, cache=FALSE}
# This is an index pairing the variable names with the order we want them
# to be shown in in the table.
tracts_data_index <- tibble(
  variable =  c(
    "obesity", "physact", "income",
    "density", "fulltime", "college", "single",
    "youth", "young_adults",  "seniors", "black", "asian", "hispanic",
    "other"),
  source = c("CDC 500 Cities", "CDC 500 Cities",
             rep("ACS", 12))
) %>%
  mutate(order = row_number())

# Get the tract SE / health data, calculate group statistics, and get it
# in the right order
tract_summary <- tracts %>%
  st_set_geometry(NULL) %>%
  tbl_df() %>%
  dplyr::select(tracts_data_index$variable) %>%
  gather(variable, value) %>%
  group_by(variable) %>%
  summarise(
    min = min(value),
    `msd` = paste(median_iqr(value)),
    max = max(value)
  ) %>%
  inner_join(tracts_data_index) %>%
  arrange(order) %>%
  mutate(
    variable = c(
      "Obesity: Share of adults over 18 who are obese",
      "Physical Activity: Share of adults over 18 who engage in some leisure-time physical activity",
      "Income: Median tract income",
      "Density: Households per square kilometer",
      "Fulltime: Share of adults over 18 with full-time work",
      "College: Share of adults over 24 with a college degree",
      "Single: Share of adults over 18 living alone or in a non-partnership household",
      "Youth: Share of population under 18",
      "Young adults: Share of population between 18 and 34",
      "Seniors: Share of population who are 65 and over",
      "Black: Share of population who is black",
      "Asian: Share of population who is Asian",
      "Hispanic: Share of population who is Hispanic",
      "Other: Share of population who belong to other minority groups")
  ) %>%
  dplyr::select(-order)

tract_rows <- nrow(tract_summary)

park_summary <- parks %>%
  st_set_geometry(NULL) %>%
  # TODO: This should be tweets
  dplyr::select(acres, courts, playgrounds, trails, tweet_count) %>% 
  gather(variable, value) %>%
  group_by(variable) %>%
  summarise(
    min = min(value),
    `msd` = paste(median_iqr(value)),
    max = max(value)
  ) %>%
  mutate(
    variable = c("Size: Park size in acres",
                 "Courts: Presence of sport courts / ball fields",
                 "Playgrounds: Presence of playgrounds",
                 "Trails: Presence of trails",
                 "Tweets: Tweets emanating from park in September 2014"),
    source = c(rep("NYC", 4), "Twitter API")
    )

bind_rows(tract_summary, park_summary) %>%
  separate(variable, c("var", "desc"), sep = ": ") %>%
  mutate(
    msd = str_replace(msd, " \\(", "\\\\\\\\ \\("),
    msd = str_c("\\makecell{", msd, "}")
  ) %>%
  kable(escape = FALSE, digits = 2, format = "latex",
        table.envir = "table*", booktabs = TRUE,
        caption = "Descriptive Statistics of Tract and Park Variables",
        col.names = c(" ", "Description", "Minimum", "Median (IQR)", "Maximum", "Source")) %>%
  kable_styling(latex_options = "scale_down") %>%
  column_spec(2, width = "2in") %>%
  group_rows(paste("Tract Variables, N =", nrow(tracts)), 1, tract_rows) %>%
  group_rows(paste("Park Variables, N =", nrow(parks)), tract_rows + 1, tract_rows + nrow(park_summary))
```



## Model 
\label{subsec:model}

We predict the rate of physical activity ($y$) in a census tract as a
function of the tract's sociodemographic characteristics and choice-based
accessibility to parks with the following  specification:
\begin{equation}\label{eq:themodel}
 y_i = f(\vect{x}_i, C_i, g(W, \vect{y}, \vect{X}))
\end{equation}
where $\vect{x}_i$ is a vector of tract-level sociodemographic attributes,
$C_i$ is a choice-based accessibility measure for the tract computed from
Equation \ref{eq:cs} and $g(W, \vect{y}, X)$ represents the impact of spatial
neighborhood effects on the outcome.

Using tracts as the unit of observation in the model allows us to consider the
availability of parks at a sub-metropolitan level, but it also introduces the problem
of spatial correlation of unobserved errors and spatial dependence in the data
generating process [@Cliff1970; @Anselin1980]. Spatial correlation occurs
when spatially-distributed unobservable or missing variables (school quality,
neighborhood prestige, etc.) influence the modeled outcomes. Spatial dependence
occurs when the outcome *depends* on the outcome in neighboring areas; for
example, seeing neighbors exercising may encourage exercise. The relaxed spatial
Durbin model (SDM) proposed by @Burridge1981 controls for both processes:
\begin{equation}\label{eq:sdm}
  \vect{y}= \rho W\vect{y} + X\vect{\beta} + WX\vect{\gamma} + \vect{\epsilon}
\end{equation}
where $\rho$ is the estimated spatial correlation between the outcome variables
$y_i, y_j, \ldots  \in \vect{y}$, $W$ is a matrix specifying the spatial
relationship between observations $i$ and $j$ in the dataset,
$\vect{\beta}$ is a vector of estimable coefficients relating the attributes $X$
of an observation to its outcome, and $\vect{\gamma}$ is a vector of
estimated coefficients relating the attributes of an observation's *neighbors*
to its outcome. If only correlation and not dependence exists in the data
generating process, it may  be more efficient to estimate a restricted
spatial autoregressive model and therefore
save degrees of freedom in the estimation. @tra_autoregression_paper
illustrate a decision algorithm originally proposed by @Florax2003 which uses
a likelihood ratio test of the SDM against the restricted spatial error model (SEM):
\begin{align}
{\boldsymbol y} &= X{\boldsymbol \beta} + {\boldsymbol u},  {\boldsymbol u} = \gamma W {\boldsymbol u} +
{\boldsymbol \epsilon} \label{eq:SEM}\\
{\boldsymbol y} &= X{\boldsymbol \beta} + (I-\gamma W)^{-1}{\boldsymbol \epsilon}\notag
\end{align}
If the likelihood ratio test rejects that the models are equivalent, then the SDM
should be used because its estimates are unbiased. One consequence of using the SDM
is that the estimated coefficients do not represent the  

The researcher must assert $W$ *a priori*, though there are several
common specifications \citep{Dubin1998}. In this study we specify that tracts
with population-weighted centroids located within 1.8 Euclidean miles of each
other are considered neighbors, and the strength of the relationship is the
row-standardized (each row sums to 1) inverse of the distance between them:
\begin{equation}\label{eq:weightsrule}
W_{ij} = \begin{cases}
  {\frac{d_{ij}^{-1}} {\sum_{k=1}^n d_{ik}^{-1}}} & \text{for $d_{ij} \leq 1.8$ miles}\\
  0  & \text{for $d_{ij} > 1.8$ miles}
  \end{cases}
\end{equation}

```{r weights, warning = FALSE}
# This are developed in 03_logsums.Rmd
distances <- read_rds(here("data/distances.rds"))
W <- read_rds(here("data/W.rds"))
trMC <- trW(as(W, "CsparseMatrix"), type="MC")
```
This is a similar specification to that used by  \citet{tra_autoregression_paper}
in a study of home prices in Atlanta. We selected this specification after
comparing overall model likelihoods between it and a specification where tracts
sharing a boundary are considered neighbors.

## Accessibility 
\label{subsec:accessibility}

Extant destination choice models used in practice typically use travel time in
minutes as a travel impedance term, and employment by sector as a size term or
attraction component. Typically such models handle recreational trips together with
other non-work and non-school trips, and we therefore can find no previously
estimated coefficients using size terms relevant to park acreage or amenities.
Given this gap in the literature, we instead assert plausible values for
$\beta$ to generate the choice-based accessibility statistic used in this study.

The functional form of the accessibility statistic is adapted from
Equation \ref{eq:v_sd} by log-transforming first the distance (Euclidean, in miles)
between the park boundaries and population-weighted tract centroids and second the
park size (in acres).
\begin{equation}\label{eq:v_logged}
V_{ij} = \log(size_j)\lambda_s + \log(distance_{ij})\lambda_d
\end{equation}
We then calculate the consumer surplus $C$  (log-sum of the exponentiated
utilities) for each tract and standardize the result,
\begin{equation}\label{eq:v_logged}
C_i' = \frac{C_i - \bar{\vect{C}}}{sd(\vect{C})}
\end{equation}

We initially estimated potential values of the $\lambda$ coefficients by
applying a limited-memory, box-constrained optimization algorithm
maximizing the model log-likehood.\footnote{Using the {\texttt nlm} function
in R [@RCoreTeam2018]} The constraint required that the coefficient on
size be positive and that on distance negative; all else equal, we assert people
will prefer to use larger parks and parks that are nearer to their residence. This
resulted in $\lambda_d = -1.98$  and $\lambda_s = 0.000$. Though the ``most
likely'' parameter value, in this case it was unsatisfactory for two reasons:
first, our prior assertion that park size is a valuable attribute of a park, and
second the observation that with these maximum likelihood values tracts
bordering on the corners of Central Park were measured with poor overall
accessibility, an unacceptably unintuitive result. We subsequently adjusted the
$\lambda_s$ value to $0.3$; the relative ratio of the coefficients implies that
all else equal, an individual would travel 6.6 times further to access a park
twice as large.


Figure \ref{fig:map} shows a map of New York City with each tract highlighted
based on its relative accessibility score. The most continuous region of good park
access is in upper Manhattan and the Bronx, bracketed by Central Park and the
Bronx River. Conversely, some of the poorest-accessibility areas are in
Brooklyn tracts not immediately adjacent to Prospect Park. Because the
accessibility statistic is normalized, the worst values are slightly below $-3$,
the best somewhere above $3$.


```{r map, fig.env="figure*", echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.cap="Normalized choice-based park accessibility log-sum values in New York City."}
library(ggmap)
map <- ggmap(
  get_stamenmap(unname(st_bbox(tracts)), 
                maptype = "toner-lite"), extent = "device")
map +
  geom_sf(data = tracts, inherit.aes=FALSE,
          aes(fill = cut(access_ls, breaks = c(-Inf, -3, -2, -1, 0, 2, 3, Inf))),
          alpha = 0.8, color = NA) +
  scale_fill_brewer("Normalized Log-sum", palette = "GnBu")
```

## Results

```{r results}
pa_models <- read_rds(here("data/pa_models.rds"))
obesity_models <- read_rds(here("data/obesity_models.rds"))
```

```{r termnames}
names_for_terms <- c(
  "(Intercept)", "log(Density)", "log(Income)",
  "Fulltime", "College-educated",  "Single Adults",
  "Youth (0-17)", "Young adults (18-34)",
  "Seniors (65+)", "Black population share",
  "Asian population share",
  "Hispanic population share", "Other Minorities",
  "$\\gamma$: spatial correlation","Physical activity",
  "Logsum", "Tweets", "Attributes", "10-min walk"
)

termnames <- tibble(
  name = names_for_terms,
  term = c("Intercept", "log(density)", "log(income)", "fulltime", "college",
           "single", "youth", "young_adults", "seniors", "black", "asian", 
           "hispanic", "other",
           "lag", "physact", "access_ls", "tweets_ls", "multi_ls", "walk_10TRUE"))
```

We estimated SDM and SEM models regressing the physical activity rate
against the base covariates (without any accessibility component). A likelihood
ratio test failed rejected that the SDM is different from the SEM, so we use
the SDM specification only going forward. Recall that in an SDM the effect of a
covariate on the dependent variable is *not* the estimated coefficient. As a
consequence of this Table \ref{tab:pa_models} presents the estimated
coefficients for the base model with the socioecononomic controls and
no representation of park accessibility. For the most part the coefficients are
highly significant and of the expected sign. Tracts with higher shares
of full-time workers, college-educated adults, young adults, and high-income
households all show a greater share of individuals engaging in regular physical
activity. Conversely, tracts with greater population density and a greater
share of single adults, children, seniors, minorities of all types, and low
income households all have lower modeled rates of physical activity.

To this base model we add a measure of park accessibility considered in two
ways: a logsum-based statistic as described earlier and a more typical quarter-mile
binary access buffer. The estimated coefficients for these models are also
given in Table \ref{tab:pa_models}. In both accessibility specifications,
the access measure is positively correlated with physical activity rates; all else
equal, people in tracts with higher accessibility to parks have higher estimated
rates of physical activity. Though the estimated coefficient on the quarter-mile
buffer appears to have a larger absolute value than the logsum-based statistic,
this masks a key difference in the specification worth further comment.
The logsum allows for variation in access unavailable in the buffer-based method.
In our dataset almost every tract
(`r round(sum(tracts$access_025) / nrow(tracts) * 100, 0)`%) is within
a quarter-mile of *some* kind of park, but this park may not be large enough
to support physical activities all individuals would like to participate in.
Because tracts with only adequate park access are grouped with tracts having
excellent park access, the overall effect is comparatively unclear. The logsum-based
coefficient, on the other hand, suggests going from the least-accessible ($-3.0$)
to the most-accessible ($3.0$) tract implies roughly a $1.1$ percentage point
rise in estimated physical activity rates.

```{r effectsplot, results = "asis"}
texreg(pa_models[c("Base", "Logsum", "10-min walk")], digits = 4,
       booktabs = TRUE, use.packages = FALSE, 
       caption = "SEM Coefficients Predicting Physical Activity Rates",
       caption.above = TRUE,
       custom.note = "%stars. Standard errors in parentheses.")
```
\end{table*}




We now consider the impacts of a model where the dependent variable is
the *obesity* rate, and physical activity becomes an independent covariate
alongside the controlling variables and the accessibility metric.
Table \ref{tab:obese_models} again presents the base model without
any accessiblity measure, and models with accessibility captured through a
logsum and through a 1/4 mile buffer. As in the
physical activity models, the coefficients are typically significant and
of the expected sign, with a few exceptions: in this case an increasing hispanic
population share  and an increasing share of children have no significant
relationship on obesity. Unlike in the physical activity models however,
the inference on the accessibility statistic *does* change between
specifications. The 1/4 mile buffer neither substantially improves the model
fit nor is its term significant at any threshold, but the accessibility logsum
indicates a significant negative correlation: moving from the least to most
accessible tract decreases the expected obesity rate by approximately half a
percentage point.



```{r obesemodelchart, results="asis"}
texreg(obesity_models, digits = 4, 
       booktabs = TRUE, use.packages = FALSE, 
       caption = "SEM Coefficients Predicting Obesity Rates",
       caption.above = TRUE,
       custom.note = "%stars. Standard errors in parentheses.")
```

\subsection{Additional Amenities}

As discussed when presenting the choice-based accessibility metric above,
a key benefit it offers is the natural way in which park amenities beyond simply size
can be accommodated. For example, the number of tweets emanating from a park
might be seen as a measure of the park's popularity, or a proxy for its use. It
may be construed that parks with high popularity or use contain other amenities
that encourage residents to use the park space, raising physical activity rates
and lowering obesity rates beyond what would be expected with the park's size
and proximity alone. It is reasonable to question whether park size and Twitter
activity are highly correlated, and thus would result in a double-counting of
park size in the accessibility statistic. In our data the number of tweets is
positively but only loosely correlated  with park
size ($\\rho = $),
indicating that the Twitter data in fact could provide new information or proxy
for the desired amenities.

To test this hypothesis, we can use the twitter data described earlier and
suggested in Equation \ref{eq:v_sdt}. In this application, we again
log-transform all the components of utility:
\begin{equation}\label{eq:v_logsdt}
V_{ij} = \log(size_j)\lambda_s + \log(distance_{ij})\lambda_d + g(tweets_j)\lambda_t
\end{equation}
where $g(x)$ is a Yeo-Johnson transformation to preserve cases where $log(0)$
would be otherwise undefined \citep{Yeo2000}. As before, there is little
information by which to determine the values of $\lambda$ in this utility
specification; we use the previous values with the addition of $\lambda_t =
0.1$. The ratio between the utility coefficients implies residents would travel
almost twenty times as far or use a park three times smaller if it had twice the
Twitter activity, all else equal.

Figure \ref{fig:tweets_map} shows the percent change in the normalized log
sum statistic after including the twitter information. There is not a recognizable
overarching pattern to the changes, though tracts immediately surrounding
Prospect Park and Central Park gain even more relative accessibility, and tracts
in midtown Manhattan, Queens, and southern Bronx tend to lose some.

```{r tweets_map, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.cap="Percent change in relative accessibility to parks when accounting for Twitter activity."}
map +
  geom_sf(data = tracts %>%
            mutate(diff = tweets_ls - access_ls,
                   pct_diff = diff / access_ls * 100),
          inherit.aes=FALSE,
          aes(fill = cut(pct_diff, quantile(pct_diff,
                                            probs = seq(0, 1, by = 0.1)))),
          alpha = 0.8, color = NA) +
  scale_fill_brewer("Percent Change", palette = "PiYG")
```


Table \ref{tab:tweets_coefs} presents the estimated coefficients relating
accessibility to obesity and physical activity rates, with and without the
inclusion of Twitter data. In both cases, including twitter information in
the accessibility statistic tempers the strength of the relationship with the
dependent variable and modestly widens the standard error without substantively
affecting the significance of the test statistics or the interpretation.
From this limited example, we cannot say whether Twitter activity enhances the
ability of parks to increase physical activity rates or lower obesity rates,
though further investigation of this and other park amenities would certainly
be warranted.


# Limitations and Future Research Direction

We readily acknowledge limitations in this study. As in any study conducted
with areal data, we are at risk of falling victim to the ecological inference
fallacy, where aggregate statistics mask or contradict disaggregated or
individual-level trends. A large-sample survey of individuals in New York City,
including measured physical activities and obesity would always be preferable
to the tract-level data used in this study. It would also be
preferrable to have obtained the $\vect{\lambda}$ accessibility utility coefficients
from a high-quality survey of leisure destination choice rather than
asserting them to match our prior expectations of what constitutes quality
park access. An ideal survey to address the question would incorporate both sets
of questions: physical activity and health data on one hand and park use (including
which parks were used and how frequently) on the other. As no such dataset
exists to our knowledge, this tract-level aggregate analysis with asserted
utility coefficients is the possibility that remains.

This paper presents a holistic accessibility statistic that could, in theory,
accommodate many attributes of the destination parks as well as the people who
might use them. As an illustration: the park-going population could be separated
into at least four delineated clusters, each preferring different amenities of a
park:
\begin{itemize}
 \item{runners and cyclists: long, interesting trail systems}
 \item{sports players: soccer fields, basketball courts, or baseball diamonds}
 \item{families with small children: water features and playgrounds}
 \item{casual users: water features, gardens, performances, etc.}
\end{itemize}
An analyst could then compute the accessibility logsum for each cluster with
different utility values for each park's amenities, and obtain a measure of
a neighborhood's accessibility to park features that its residents most care
about. In this paper, we proceed only incrementally beyond this theory by
adding Twitter activity as an element of a park's attractiveness above and
beyond its size. Exploring additional amenities or market segmentation strategies
could provide a more definitive understanding of the relationship between park
accessibility and health outcomes.

As a travel impedance measure, we used the Euclidean distance between each tract's
population-weighted centroid and the nearest point on the edge of a park's border.
Euclidean distances have well-rehearsed limitations regarding their fidelity with
the underlying infrastructure network, etc. Network-based distances can also
suffer from challenges when applied to multimodal problems; these challenges are
exacerbated when the non-highway mode share is high, as is likely when considering
access to parks. A better metric of travel impedance may be a mode choice model
logsum, which weights all travel alternatives against each other.

Finally, this study is primarily focused on the hypothesis that accessibility to
parks encourages physical activity, which in turn reduces obesity. There are a
multitude of other hypotheses that might be proposed and tested with the basic
methodology we have employed here, or competing explanations for the outcomes we
have observed. It is distinctly possible, for instance, that individuals who wish
to exercise regularly in parks choose to live near them. Given that the CDC models
generating the obesity and physical activity estimates presumably include variables
likely to influence such preferences (income, etc.), our investigation cannot
isolate the preferences from the effect. Controlling for such a
self-selection effect would be necessary to isolate the exogenous
impacts of park access on obesity or other health outcomes. And regarding these
other variables: this study did not consider potential relationships between park
access and hospitalization rates, life expectancy, respiratory disease,
mental health, or any number of potential beneficial outcomes hypothesized
or explored in the existing literature. Exploring these connections and their
underlying mechanisms should be a priority as city planners and urban architects
attempt to improve the quality of life of urban residents in the future.

# Conclusion

Increased physical activity and decreased obesity rates are critical measures of
improvement in public health. Although many have theorized the link between park
accessibility and these metrics, previous literature has produced mixed
findings, perhaps owing to the range of variables modeled and the coarse spatial
scale of the analyses. Using New York City as a case study, we presented a
holistic and flexible measurement for
park accessibility that compares the continuous distance to all parks in the
region, weighted against the size of the park and its other amenities. In terms
of physical activity, we found a positive relationship where the least
park-accessible tracts have an expected physical activity participation rate
roughly one percentage point lower than the most accessible tracts. In terms of
obesity, we also found a strong and significant correlation between increased
park accessibility and decreased obesity rates, controlling for
spatial correlation, socioeconomic attributes, and physical activity rates.


In both cases a traditional, buffer-based analysis estimated a weaker relationship
with greater uncertainty. Buffer analyses are relatively common in the literature,
perhaps because of the widespread availability of GIS software. In spite of their
widespread use, they are relatively limited in terms of both their theoretical
underpinnings and their flexibility to accommodate attributes of parks beyond
their proximity. And even proximity may not be adequately handled, as the buffer
distance may be arbitrarily asserted by the researcher. The model we develop and
apply in this paper extends buffer driven models to contribute a more
comprehensive and flexible approach for measuring the impact of park access on
health outcomes. Adopting choice-based accessibilities of the kind used in this
study will allow researchers to
encompass the full range of park amenities in their accessibility analyses.
This will in turn enable planners to consider how multiple attributes of a park
--- from its location to its size to its amenities and beyond --- benefit the health
of the community the park serves.

## Acknowledgements
This project was funded by the Speedwell Foundation. We are grateful to Rachel
Samuels for helping to assemble the geolocated Twitter data.

# References {#references .unnumbered}